Apropos Replacement: Development of a full text search tool for man pages \newline (Initial Outline)

### 2011-12-26 paper.tex
Minor fixes

\documentclass[titlepage, a4paper, 12pt]{article}
\usepackage{listings}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[colorlinks=false]{hyperref}
\lstset{ %
language=sh,
basicstyle=\footnotesize,
numbers=none,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname
}
\title{Apropos Replacement \newline
(Initial Outline)}
\author{Abhinav Upadhyay $<$er.abhinav.upadhyay@gmail.com$>$ \and 
Joerg Sonnenberger $<$joerg@NetBSD.org$>$}
\begin{document}
\maketitle
\section{Abstract}
Unix like operating systems have withstood the test of time for 40 odd years for
several technical and philosophical reasons. One of those reasons is the
quality documentation they ship in the form of man pages, which are a single
source of consultation for most of our routine work. However, so far, there has been a lack of
good search tool to accompany this documentation, which would increase their
usefulness manifolds.

@@

In this paper, the limitations of the traditional implementation of
\textit{apropos} are discussed along with brief details of how these were overcome as part of this.
Additionally, we compare and contrast our approach with other modern implementations available out there.

@@

This simplified design of \textit{apropos(1)} resulted in limited utility and
usability as the searches are limited to the keywords in the NAME section of the
 keywords. Only if
a user knows the exact keywords do they get the appropriate results; otherwise most of
the times the searches are full of irrelavant results or possibly a dead end. In the modern
computing world where hard problems like searching the World Wide Web have been
solved [3] to a sufficient degree, it makes perfect sense to leverage the advancement in technology 
in order to solve the problem. This problem is made even simpler considering that the man
pages consist of structured data, making the problem somewhat less complex.

@@

\section{Limitations of Conventional \textit{apropos(1)}}
In this section results of some sample query searches from
the classical version of \textit{apropos} are shown and their shortcomings are
analyzed.

\subsection{Lack of support for free form queries}
The conventional \textit{apropos} is limited to the keywords used in the NAME
section
of the man pages. If the user specifies keywords relevant to the page which he
is looking but if those keywords are not used in the NAME section,
then \textit{apropos} would not return
anything relevant. So, in essence, the user cannot use free form queries like
those supported by modern search applications.
\lstset{language = sh, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos ``add new user''
add new user: nothing appropriate

@@

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it has no
natural language support. For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query \textit{``make directories''}, it fails when the keyword \textit{``directory''} is
used in place of \textit{``directories''} even though the two words are based on
same root word.
\lstset{language = sh, caption={No support for stems or word with same roots},}
\begin{lstlisting}
$ apropos ``make directories''
mkdir (1) - make directories
$ apropos ``make directory''
make directory: nothing appropriate

@@

\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ \textit{apropos(1)} power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_class_display_register, pmf_system_suspend, pmf_system_resume,
pmf_system_shutdown, pmf_event_register, pmf_event_deregister,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
\subsection{Other Problems}
Apart from the search related problems there are a few issues related to
the way man pages are handled in NetBSD. The different aliases of the man pages
are stored on the filesystem in the from of hard (or soft) links and these
have to be mentioned in the makefiles explicitly using the MLINKS mechanism. This
approach works fine but it is a mess from maintenance point of view.
It should be possible to fix this by utilizing the index already built and
maintained by \textit{apropos} and related tools, but yet again, the
simplistic implementation of \textit{apropos} does not leave any room for
improvement.

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems as
part of this project.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the keywords defined in the NAME section of the man
pages as the keywords used throughout the body of the page also play a role.
e.g. queries like \textit{``installing new software package''}
will now produce relevant results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation, the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This enables support for more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

Similarly along with the inverted index, a dictionary of the keywords frequently
occurring in the corpus of man pages can also be built and used to support
spelling suggestion.

In this implementation, both word stemming and spelling correction in our implementation are used to make the search experience more smooth.
\end{description}
\begin{description}
\item[Bookkeeping of man page metadata] \hfill \\
In this implementation additional metadata related to the man pages, for example
an md5 hash, the device id, inode number and modification time
of the man page files are indexed and stored so
as to support fast and hassle free update of the index as new man pages are
installed or the old ones are modified/updated.

@@

The two main tasks that are needed to be performed to implement this project are
parsing of the man pages
and building an index from the data. We have used \textit{libmandoc} [6]
and \textit{sqlite} [7] have for these the two tasks.

@@

interface to parse and build an AST (Abstract Syntax Tree) of the man page. It also provides
an interface for traversing that tree in order to extract the data from nodes
which are of our interest.

@@

relatively small and easy to use C library interface. One of the main reason

@@

of the application. Besides that another advantage of \textit{sqlite} is the
RDBMS support which makes it very easy to store additional metadata in the form

@@

Due to space constraints it is not possibile to go into enough implementation details, we discuss 
the most important components of this project  in the section below:

@@

is to develop different interfaces on top of this implementation for different

@@

in normal case for a query but in the following listings the output has been

@@

\begin{figure}[htp]
\includegraphics[scale=0.30]{/home/abhinav/development/AsiaBSDCon/spell.png}
\caption{The spell corrector and the web interface in action}
\label{}
\end{figure}
\section{Related Work}
There are at least two projects which are related to this in some way.
\begin{description}
\item[man-db] \hfill \\
\textit{man-db} [11] is a complete implementation of the man page documentation
system and it is used on a number of GNU/Linux distributions.\textit{man-db}
takes an interesting approach for indexing the man page data. Unlike the
classical \textit{apropos}, it uses a Berkley DB database but still it's index
is limited to the NAME section only. It adds an option `K' to man(1) to allow a
crude full text search but it is not very efficient nor effective.
\end{description}
\begin{description}
\item[mandocdb] \hfill \\
\textit{mandocdb} [12] is more in line with the goals of our project but it takes
a novel approach. It indexes the keywords extracted from the man pages in a
key-value store using btree(3) [13]. It also comes with it's own implementation
of \textit{apropos} which performs search using this key-value store. The
main key point of this implementation is that it exploits the semantic structure
of the man pages. 
\end{description}
\section{Future Work}
\begin{description}
\item[Work on Ranking Algorithm] \hfill \\
A ranking algorithm based on probabilistic model [14] of information
retrieval has been implemented to filter out the most relevant results and show
them at the top. It is essentially based on the Okapi BM25F algorithm [15] and
uses certain parameters whose values are usually dependent on the corpus and the
search application. For example certain weight parameters are assigned to different sections of man pages. At the moment these values have been determined
manually but one goal is to use some machine learning techniques in the coming
days to automate this.
\end{description}
\begin{description}
\item[Fix Some Loose Ends In Parsing of man pages] \hfill \\
Although the parsing routine is working fine for most purposes on the man pages
supplied with NetBSD and also on most of the man pages found in Pkgsrc. But there
are perhaps a few corner cases still to be fixed. Besides that another issue that
is to be addressed is the parsing of the escape sequences which are used
extensively in the \textit{man(7)} [16] based man pages. These issues are to be
addressed in the coming days.
\end{description}
\section{Acknowledgement}
This project has been developed as part of \textbf{Google Summer of Code 2011}
[17], so thanks to \textbf{Google} for sponsoring it. Special thanks to
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. We would like to thank \textbf{David Young} who was involved with
this project closely and offered useful help and guidance throughout. A special
thanks goes to \textbf{Thomas Klausner} who helped in writing and reviewing our
man pages. Thanks to \textbf{Petra Zeidler} for administering the GSoC program
for \textit{The NetBSD Foundation}.
\begin{thebibliography}{20}
\bibitem{1}
\emph{NetBSD manual page for apropos(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?apropos++NetBSD-5.1

@@

\bibitem{14}
Fuhr, Norbert
\emph{Probabilistic models in information
retrieval}
The Computer Journal
1992
\bibitem{15}
Zaragoza H.; Craswell N.; Taylor M.;
Saria S.; Robertson S.
\emph{Microsoft Cambridge at TREC–13: Web
and HARD tracks}
In proceedings of TREC-2004

>>>>

\documentclass[titlepage, a4paper, 12pt]{article}
\usepackage{listings}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage[colorlinks=false]{hyperref}
\lstset{ %
language=sh,
basicstyle=\footnotesize,
numbers=none,
backgroundcolor=\color{white},
showspaces=false,
showstringspaces=false,
showtabs=false,
frame=single,
tabsize=2,
captionpos=b,
breaklines=true,
breakatwhitespace=false,
title=\lstname
}
\title{Apropos Replacement: Development of a full text search tool for man pages \newline
(Initial Outline)}
\author{Abhinav Upadhyay $<$er.abhinav.upadhyay@gmail.com$>$ \and 
Joerg Sonnenberger $<$joerg@NetBSD.org$>$}
\begin{document}
\maketitle
\section{Abstract}
Unix like operating systems have withstood the test of time for 40 odd years for
several technical and philosophical reasons. One of those reasons is the
quality documentation they ship in the form of man pages, which are a single
source of consultation for most of our routine work. However, so far, there has been a lack of
good search tool to accompany this documentation, which would increase their
usefulness manifolds.

@@

In this paper, the limitations of the traditional implementation of
\textit{apropos} are discussed along with brief details of how these were overcome as part of this project.
Additionally, we compare and contrast our approach with other modern implementations available out there.

@@

This simplified design of \textit{apropos} resulted in limited utility and
usability as the searches are limited to the keywords defined in the NAME
section of the man pages. Only if a user knows the exact keywords do they get
the appropriate results; otherwise most of
the times the searches are full of irrelevant results or possibly a dead end. In the modern
computing world where hard problems like searching the World Wide Web have been
solved [3] to a sufficient degree, it makes perfect sense to leverage the
advancement in technology in order to solve the problem. This problem is made
even simpler considering that the man pages consist of structured data.

@@

\section{Limitations of Conventional \textit{apropos}}
In this section results of some sample query searches from
the classical version of \textit{apropos} are shown and their shortcomings are
analyzed.

\subsection{Lack of support for free form queries}
As noted earlier the conventional \textit{apropos} is limited to the keywords
used in the NAME section of the man pages. This means that the users do not have
a whole lot of choice for specifying keywords in their queries. Besides they
cannot search for keywords which are not usually specified in the title of a man
page, for example a query like ``EINVAL'' would return no results.
\lstset{language = sh, caption={No results for free form queries}}
\begin{lstlisting}
$ apropos ``add new user''
add new user: nothing appropriate

@@

\subsection{Lack of basic language support}
Another major limitation of the classical version of \textit{apropos} is that
it is not smart enough to provide basic natural language processing constructs
like stemming and spelling corrections.
For example in the following listing it can be seen
that while \textit{apropos} returns the correct result for the query
\textit{``make directories''}, it fails when the keyword \textit{``directory''}
is used in place of \textit{``directories''}, even though the two words are based
on same root word.
\lstset{language = sh, caption={No support for stems or word with same roots},}
\begin{lstlisting}
$ apropos ``make directories''
mkdir (1) - make directories
$ apropos ``make directory''
make directory: nothing appropriate

@@

\lstset{language = sh, caption={Unintelligible output of \textit{apropos(1)}},}
\begin{lstlisting}
$ \textit{apropos(1)} power
PCI, pci_activate, pci_bus_devorder, pci_chipset_tag_create,
pci_chipset_tag_destroy, pci_conf_read, pci_conf_write,
pci_conf_print, pci_conf_capture, pci_conf_restore, pci_find_device,
pci_get_capability, pci_mapreg_type, pci_mapreg_map, pci_mapreg_info,
pci_intr_map, pci_intr_string, pci_intr_evcnt, pci_intr_establish,
pci_intr_disestablish, pci_get_powerstate, pci_set_powerstate,
pci_vpd_read, pci_vpd_write, pci_make_tag, pci_decompose_tag,
pci_findvendor, pci_devinfo, PCI_VENDOR, PCI_PRODUCT, PCI_REVISION (9)
- Peripheral Component Interconnect
PMF, pmf_device_register, pmf_device_register1, pmf_device_deregister,
pmf_device_suspend, pmf_device_resume, pmf_device_recursive_suspend,
pmf_device_recursive_resume, pmf_device_resume_subtree,
pmf_class_network_register, pmf_class_input_register,
pmf_class_display_register, pmf_system_suspend, pmf_system_resume,
pmf_system_shutdown, pmf_event_register, pmf_event_deregister,
pmf_event_inject, pmf_set_platform, pmf_get_platform (9) - power
management and inter-driver messaging framework
acpi (4) - Advanced Configuration and Power Interface
acpipmtr (4) - ACPI Power Meter
amdpm (4) - AMD768 Power Management Controller and AMD8111 System
Management Controller
...
.
.
.
\end{lstlisting}
\subsection{Other Problems}
Apart from the search related problems there are a few issues related to
the way man pages are handled in NetBSD. The different aliases of the man pages
are stored on the filesystem in the from of hard (or soft) links and these
have to be mentioned in the makefiles explicitly using the MLINKS mechanism. This
approach works fine but it is a mess from maintenance point of view.
It should be possible to fix this by utilizing the index already built and
maintained by \textit{apropos}, but yet again, the simplistic implementation of
\textit{apropos} does not leave any room for improvement.

\section{Proposed Solution}
A very simple and straightforward solution is proposed to solve these problems as
part of this project.
The idea is to parse and index
the complete content (or at least most of the content) in the form of an
inverted index [4] and use it to build a full text search
interface. Having an index based on the complete content of the man pages solves
many of the search
related problems associated with the conventional \textit{apropos(1)} as noted
before and discussed as follows:
\begin{description}
\item[Free Form Queries] \hfill \\
The users can express their queries in more natural language form as the searches
are no longer limited to the the NAME section.
For example, queries like \textit{``installing new software package''}
now produce relevant results.
\end{description}
\begin{description}
\item[Basic Natural Language Processing Support] \hfill \\
When parsing the man pages and building the index, it is also possible to
pre-process the tokens extracted from the man pages to support some of the very
basic natural language processing functionalities. In this implementation, the
Porter stemming algorithm [5] has been used to
reduce the individual tokens extracted from the man pages to their root words.
This enables support for more flexible searches. For example both
\textit{``Installing new packages''} and \textit{``install new package''} will
return same results.

Similarly along with the inverted index, a dictionary of the keywords frequently
occurring in the corpus of man pages is also built and used to support
spelling suggestion.
\end{description}
\begin{description}
\item[Bookkeeping of man page metadata] \hfill \\
In this implementation, additional metadata related to the man pages, for example
an md5 hash, the device id, inode number and modification time
of the man page files are indexed and stored so
as to support fast and hassle free update of the index as new man pages are
installed or the old ones are modified/updated.

@@

The two main operations that are critical for this project are parsing of the
man pages and building of an index of the
data obtained from the parsing process. \textit{libmandoc} [6]
and \textit{sqlite} [7] have been used for executing these the two tasks.

@@

interface to parse and build an AST (Abstract Syntax Tree) of the man page. It
also provides an interface for traversing that tree in order to extract the data
from nodes which are of our interest.

@@

relatively small and easy to use C library interface. One of the main reasons

@@

of the application. Besides that, another advantage of \textit{sqlite} is the
RDBMS support, which makes it very easy to store additional metadata in the form

@@

Due to space constraints it is not possible to go into enough implementation
details, the most important components of this project are described in brief
in this section.

@@

is to develop different interfaces on top of it for different

@@

in normal cases for a query but in the following listings the output has been

@@

\begin{figure}[htp]
\includegraphics[scale=0.30]{/home/abhinav/development/AsiaBSDCon/spell.png}
\caption{The spell corrector and the web interface in action}
\label{}
\end{figure}
\section{Related Work}
There are at least two projects which are related to this in some way.
\begin{description}
\item[man-db] \hfill \\
\textit{man-db} [11] is a complete implementation of the man page documentation
system and it is used on a number of GNU/Linux distributions.\textit{man-db}
takes an interesting approach for indexing the man page data. Unlike the
classical \textit{apropos}, it uses a Berkley DB database but still it's index
is limited to the NAME section only. It adds an option `K' to man(1) to allow a
crude full text search but it is not very efficient nor effective.
\end{description}
\begin{description}
\item[mandocdb] \hfill \\
\textit{mandocdb} [12] is more in line with the goals of this project but it
takes a novel approach. It indexes the keywords extracted from the man pages in a
key-value store using btree(3) [13]. It also comes with it's own implementation
of \textit{apropos} which performs search using this key-value store. The
main key point of this implementation is that it exploits the semantic structure
of the man pages.
\end{description}
\section{Future Work}
\begin{description}
\item[Work on Ranking Algorithm] \hfill \\
A ranking algorithm based on probabilistic model [14] of information
retrieval has been implemented to filter out the most relevant results and show
them at the top. It is essentially based on the Okapi BM25F algorithm [15] and
uses certain parameters whose values are usually dependent on the corpus and the
search application. For example certain weight parameters are assigned to
different sections of man pages. At the moment these values have been determined
manually but one goal is to use some supervised machine learning techniques in
the order to automate this.
\end{description}
\begin{description}
\item[Fix Some Loose Ends In Parsing of man pages] \hfill \\
Although the parsing routine is working fine for most purposes on the man pages
supplied with NetBSD and also on most of the man pages found in Pkgsrc. But there
are perhaps a few corner cases still to be fixed. Besides that, another issue
that
is to be addressed is the parsing of the escape sequences which are used
extensively in the \textit{man(7)} [16] based man pages. These issues are to be
addressed in the coming days.
\end{description}
\section{Acknowledgement}
This project has been developed as part of \emph{Google Summer of Code 2011}
[17], so thanks to \textbf{Google} for sponsoring it. Special thanks to
\textbf{Kristaps Dzonsons} who is the developer of the \textit{mdocml} [18]
project, he also helped by pointing out several issues in the parsing related
code. We would like to thank \textbf{David Young} who was involved with
this project closely and offered useful help and guidance throughout. A special
thanks goes to \textbf{Thomas Klausner} who helped in writing and reviewing
man pages for this project. Thanks to \textbf{Petra Zeidler} for administering
the GSoC program for \textit{The NetBSD Foundation}.
\begin{thebibliography}{20}
\bibitem{1}
\emph{NetBSD manual page for apropos(1)}
\newline http://netbsd.gw.com/cgi-bin/man-cgi?apropos++NetBSD-5.1

@@

\bibitem{14}
Fuhr, Norbert
\emph{Probabilistic models in information
retrieval}
The Computer Journal
1992
\bibitem{15}
Zaragoza, H.; Craswell, N.; Taylor, M.;
Saria, S.; Robertson, S.
\emph{Microsoft Cambridge at TREC–13: Web
and HARD tracks}
In proceedings of TREC-2004
